Natural selection in the Health and Retirement Study

### 2023-09-16 hrs-selection.qmd
Use svyglm and create survey design objects

library(haven)
library(forcats)
library(dplyr, warn.conflicts = FALSE)
library(stringr)
library(purrr)
library(ggplot2)
library(tidyr)
library(santoku, warn.conflicts = FALSE) # masks tidyr::chop

@@

# weighting
rand <- left_join(rand, trk_weights, by = join_by(hhid == HHID, pn == PN))
rand$weight <- ifelse(! is.na(rand$MBIOWGTR), rand$MBIOWGTR, rand$NBIOWGTR)

@@

for (eth in eths) {
  r_eth <- rand |> filter(ethnicity == eth)

  form_pgs <- map(pgs, \(x) reformulate(c(x, pcs), response = "raevbrn"))
  mod_pgs[[eth]] <- map(form_pgs, 
                        \(x) lm(x, data = r_eth, weights = weight)
                        )

@@

  form_pgs_edyrs <- map(pgs, \(x) reformulate(c(x, pcs),
                                              response = "raedyrs"))
  mod_pgs_edyrs[[eth]] <- map(form_pgs_edyrs, 
                              \(x) lm(x, data = r_eth, weights = weight)
                              )

@@

  form_pgs_sibs <- map(pgs, 
                       \(x) reformulate(c(x, pcs), response = "r10sibs"))
  mod_pgs_sibs[[eth]] <- map(form_pgs_sibs,
                             \(x) lm(x, data = r_eth, 
                                     weights = weight/(1 + r10sibs)))

@@

  # the colon interaction estimates the effect separately within each group
  form_pgs_x_ed <- paste0(pgs, ":edu") |>
    map(\(x) reformulate(c(x, "edu", pcs), response = "raevbrn"))
  mod_pgs_x_ed[[eth]]  <-
    map(form_pgs_x_ed, 
        \(x) lm(x, data = r_eth, weights = weight)
        )

@@

  form_pgs_x_inc <- paste0(pgs, ":income_med") |>
    map(\(x) reformulate(c(x, "income_med", pcs), response = "raevbrn"))
  mod_pgs_x_inc[[eth]]  <-
    map(form_pgs_x_inc, 
        \(x) lm(x, data = r_eth, weights = weight)
        )

@@

  form_pgs_x_mar <- paste0(pgs, ":married") |>
    map(\(x) reformulate(c(x, "married", pcs), response = "raevbrn"))
  mod_pgs_x_mar[[eth]]  <-
    map(form_pgs_x_mar, 
        \(x) lm(x, data = r_eth, weights = weight)
        )

@@

  form_pgs_x_agefb <- paste0(pgs, ":agefb") |>
    map(\(x) reformulate(c(x, "agefb", pcs), response = "raevbrn"))
  mod_pgs_x_agefb[[eth]]  <-
    map(form_pgs_x_agefb, 
        \(x) lm(x, data = r_eth, weights = weight)
        )

@@

  cor_fert_edu_phen[[eth]] <- 
    lm(raevbrn ~ raedyrs, data = r_eth, weights = weight) |> 
    coef() |> 
    pluck("raedyrs")
}

@@

n_tests <- nrow(tidy_pgs_eth)
n_sig_bonf <- sum(tidy_pgs_eth$p.value < 0.05/n_tests)
if (n_sig_bonf != 1) stop("More/less than 1 score was Bonferroni significant")

@@

Appendix @fig-fertility shows effect sizes of PGS correlations with realised fertility among each ethnic group. Standard errors are large because of the relatively low sample sizes, and only `r n_sig_bonf` score is significant at Bonferroni-corrected p \< 0.05/`r n_tests`. But I am most concerned with looking at patterns across scores rather than judging the significance of individual scores.

@@

n_reps <- 599
cors_boot <- list()
for (eth in eths) {
  rand_short <- rand |> 
    filter(ethnicity == eth) |> 
    select(all_of(pgs), all_of(pcs), raevbrn, raedyrs, weight, r10sibs)
  cors_boot[[eth]] <-
    replicate(n_reps, {
      r_boot <- rand_short |> 
              slice_sample(prop = 1, replace = TRUE)

      coef_pgs_fert <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "raevbrn")) |> 
        map(\(x) lm(x, data = r_boot, weights = weight)) |> 
        map_dbl(\(x) coef(x)[[2]])

      coef_pgs_sibs <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "r10sibs")) |> 
        map(\(x) lm(x, data = r_boot, weights = weight/(r10sibs + 1))) |> 
        map_dbl(\(x) coef(x)[[2]])

      coef_pgs_edyrs <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "raedyrs")) |> 
        map(\(x) lm(x, data = r_boot, weights = weight)) |> 
        map_dbl(\(x) coef(x)[[2]])

@@

for (eth in eths) {}

>>>>

library(haven)
library(forcats)
library(dplyr, warn.conflicts = FALSE)
library(stringr)
library(purrr)
library(ggplot2)
library(tidyr)
library(santoku, warn.conflicts = FALSE) # masks tidyr::chop
library(survey)

@@

# weighting
rand <- left_join(rand, trk_weights, by = join_by(hhid == HHID, pn == PN))
rand$weight <- ifelse(! is.na(rand$MBIOWGTR), rand$MBIOWGTR, rand$NBIOWGTR)
# compensate for fecund parents more likely to have child in survey:
rand$parent_weight <- rand$weight / (rand$r10livsib + 1)

rand <- rand |> 
  select(ethnicity, all_of(pgs), all_of(pcs), weight, raevbrn, raedyrs,
         r10sibs, edu, income_med, married, agefb, rabyear,
         raehsamp, raestrat, parent_weight)

r_parents <- svydesign(
    id = ~ raehsamp,
		strata = ~ raestrat,
		weights = ~ parent_weight, 
		nest = TRUE,
		data = rand |> drop_na(parent_weight) 
		)

rand <- svydesign(
    id = ~ raehsamp,
		strata = ~ raestrat,
		weights = ~ weight, 
		nest = TRUE,
		data = rand |> drop_na(weight) 
		)

@@

for (eth in eths) {
  r_eth <- rand |> subset(ethnicity == eth)
  r_parents_eth <- r_parents |> subset(ethnicity == eth)

  form_pgs <- map(pgs, \(x) reformulate(c(x, pcs), response = "raevbrn"))
  mod_pgs[[eth]] <- map(form_pgs, 
                        \(x) svyglm(x, r_eth)
                        )

@@

  form_pgs_edyrs <- map(pgs, \(x) reformulate(c(x, pcs),
                                              response = "raedyrs"))
  mod_pgs_edyrs[[eth]] <- map(form_pgs_edyrs, 
                              \(x) svyglm(x, r_eth)
                              )

@@

  form_pgs_sibs <- map(pgs, 
                       \(x) reformulate(c(x, pcs), response = "r10sibs"))
  mod_pgs_sibs[[eth]] <- map(form_pgs_sibs,
                             \(x) svyglm(x, r_parents_eth))

@@

  # the colon interaction estimates the effect separately within each group
  form_pgs_x_ed <- paste0(pgs, ":edu") |>
    map(\(x) reformulate(c(x, "edu", pcs), response = "raevbrn"))
  mod_pgs_x_ed[[eth]]  <-
    map(form_pgs_x_ed, 
        \(x) svyglm(x, r_eth)
        )

@@

  form_pgs_x_inc <- paste0(pgs, ":income_med") |>
    map(\(x) reformulate(c(x, "income_med", pcs), response = "raevbrn"))
  mod_pgs_x_inc[[eth]]  <-
    map(form_pgs_x_inc, 
        \(x) svyglm(x, r_eth)
        )

@@

  form_pgs_x_mar <- paste0(pgs, ":married") |>
    map(\(x) reformulate(c(x, "married", pcs), response = "raevbrn"))
  mod_pgs_x_mar[[eth]]  <-
    map(form_pgs_x_mar, 
        \(x) svyglm(x, r_eth)
        )

@@

  form_pgs_x_agefb <- paste0(pgs, ":agefb") |>
    map(\(x) reformulate(c(x, "agefb", pcs), response = "raevbrn"))
  mod_pgs_x_agefb[[eth]]  <-
    map(form_pgs_x_agefb, 
        \(x) svyglm(x, r_eth)
        )

@@

  cor_fert_edu_phen[[eth]] <- 
    svyglm(raevbrn ~ raedyrs, r_eth) |> 
    coef() |> 
    pluck("raedyrs")
}

@@

n_tests <- nrow(tidy_pgs_eth)
n_sig_bonf <- sum(tidy_pgs_eth$p.value < 0.05/n_tests)
if (n_sig_bonf != 0) stop("More/less than 0 score was Bonferroni significant")

@@

Appendix @fig-fertility shows effect sizes of PGS correlations with realised fertility among each ethnic group. Standard errors are large because of the relatively low sample sizes, and no scores are significant at Bonferroni-corrected p \< 0.05/`r n_tests`. But I am most concerned with looking at patterns across scores rather than judging the significance of individual scores.

@@

n_reps <- 599
cors_boot <- list()
for (eth in eths) {
  rand_short <- rand |> 
    subset(ethnicity == eth)
  cors_boot[[eth]] <-
    replicate(n_reps, {
      # survey package automatically does the right thing
      # possibly I should be using survey::bootweights
      # or survey::withReplicates and a function
      r_boot <- rand_short[sample(nrow(rand_short), replace = TRUE), ]

      coef_pgs_fert <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "raevbrn")) |> 
        map(\(x) svyglm(x, r_boot)) |> 
        map_dbl(\(x) coef(x)[[2]])

      coef_pgs_sibs <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "r10sibs")) |> 
        # need to add , weights = weight/(r10sibs + 1)
        map(\(x) svyglm(x, r_boot)) |> 
        map_dbl(\(x) coef(x)[[2]])

      coef_pgs_edyrs <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "raedyrs")) |> 
        map(\(x) svyglm(x, r_boot)) |> 
        map_dbl(\(x) coef(x)[[2]])

@@
